---
title: "G2M"
author: "Francis Gichere"
date: "13/03/2021"
output: html_document
---



### Loading libraries to be used
```{r, message=FALSE, warning=FALSE}
### Loading the required libraries
library(tidyverse)
library(lubridate)
library(tidyr)
library(dlookr)
library(stringr)
library(ggpubr)
library(knitr)
library(pastecs)
library(reshape2)
library(ggcorrplot)
```

#loading the datasets

```{r}
#Cab Data
cab<-read_csv("Cab_Data.csv")
#City data
city<-read_csv("City.csv")
#Customer ID dataset
customerID<-read_csv("Customer_ID.csv")
#Transaction Id data
transactionID<-read_csv("Transaction_ID.csv")

```

### Checking the structure of the data sets and data types
#### 1. Cab Data
```{r}
# 1. Cab Data
glimpse(cab)
#Changing Date of travel from character to date data type
cab$`Date of Travel`<-dmy(cab$`Date of Travel`)
#saving the new dataset
write_csv(cab, "cab.csv")
#Loading the saved dataset
cab_data<-read_csv("cab.csv")
head(cab_data)
```

#### 2. City data

```{r}
glimpse(city)
head(city)
```

#### 3. Customer ID data

```{r}
glimpse(customerID)
head(customerID)
```

#### 4. Transactions data

```{r}
glimpse(transactionID)
head(transactionID)
```

### Combining data sets
Combining Transaction ID data with cab data set
```{r}
cab_trans<-left_join(transactionID, cab_data, by="Transaction_ID")
#checking the structure of the new data set
glimpse(cab_trans)
#taking a look at the first few rows of the data
head(cab_trans)
#saving the new data set to the working directory
write_csv(cab_trans, "cab_trans.csv")
```

Joining customer ID dataset with TransactionId and Cab data

```{r}
master_data<-left_join(cab_trans, customerID, by="Customer_ID")
glimpse(master_data)
## Dropping the rows with missing values
master_clean<-master_data %>% drop_na()
glimpse(master_clean)
#Adding a calculated column for profit margin
master_clean$Profit_margin<-(master_clean$`Price Charged` - master_clean$`Cost of Trip`)
glimpse(master_clean)
#Writing the final data set to be used in the analysis

write_csv(master_clean, "g2m.csv")
```

### Exploratory Data Analysis

```{r}
final_data<-read_csv("g2m.csv")
head(final_data)
glimpse(final_data)
```

####Defining Gender and Company as factors
```{r}
final_data$Company<-as.factor(final_data$Company)
final_data$Gender<-as.factor(final_data$Gender)
final_data$Payment_Mode<-as.factor(final_data$Payment_Mode)
### Creating age groups
final_data$Age_Group<-cut(final_data$Age, c(17,25,32,39,46,53,60,65),
                          labels = c("18-25",
                                     "26-32",
                                     "33-39",
                                     "40-46",
                                     "47-53", 
                                     "54-60",
                                     "61+"))
```


## Plots
#### Distributions

```{r}
# Kilometer traveled 
ggplot(final_data, mapping = aes(x=`KM Travelled`))+
  geom_histogram(mapping = aes(color=Company), bins = 30)
```

```{r}
# Price charged
ggplot(final_data, mapping = aes(x=`Price Charged`))+
  geom_histogram(mapping = aes(color=Company), bins = 30)
```

```{r}
# Cost of trip
ggplot(final_data, mapping = aes(x=`Cost of Trip`))+
  geom_histogram(mapping = aes(color=Company), bins = 30)
```

```{r}
# Income (USD/Month)
ggplot(final_data, mapping = aes(x=`Income (USD/Month)`))+
  geom_histogram(mapping = aes(color=Gender), bins = 30)
```

```{r}
# Profit Margin
ggplot(final_data, mapping = aes(x=Profit_margin))+
  geom_histogram(mapping = aes(color=Company), bins = 30)
```

#### Checking Associations between variables
```{r}

```

